Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-22 draft/Results.tex
sss

%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}

@@

 \begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
	\scalebox{0.7}{
		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
 Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline
ellipse	&155	&9	&0	&0	&0	&5	&0	&0	&0	&0	&0	&6	&0\\ \hline
heart	&0	&109	&1	&1	&25	&0	&0	&1	&0	&0	&2	&1	&1\\ \hline
trapezoid	&0	&0	&103	&5	&2	&0	&0	&2	&0	&0	&34	&0	&0\\ \hline
pentagon	&0	&0	&20	&121	&0	&20	&6	&2	&0	&0	&1	&0	&0\\ \hline
arch	&0	&1	&13	&0	&86	&0	&0	&2	&3	&0	&2	&0	&0\\ \hline
hexagon	&0	&0	&0	&0	&0	&98	&0	&0	&0	&0	&0	&0	&0\\ \hline
square	&0	&0	&1	&0	&2	&0	&107	&14	&0	&0	&8	&1	&0\\ \hline
triangle	&0	&0	&7	&0	&1	&0	&0	&89	&0	&0	&12	&4	&0\\ \hline
cube	&0	&0	&0	&0	&0	&0	&0	&0	&147	&0	&0	&0	&0\\ \hline
cylinder	&0	&0	&0	&6	&2	&8	&16	&0	&4	&156	&1	&10	&0\\ \hline
parallelogram	&0	&0	&4	&0	&0	&0	&0	&0	&0	&0	&91	&0	&0\\ \hline
moon	&0	&3	&3	&2	&17	&0	&9	&17	&0	&0	&2	&133	&11\\ \hline
callout	&1	&34	&0	&22	&21	&22	&16	&25	&0	&0	&0	&0	&143\\ \hline
		\end{tabular}
			}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}

@@

\subsubsection{Features Comparisons}
\label{sec:featuresComparisions}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAllHS} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.  Figure \ref{fig:testFeaturesAllMD} shows the similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset. The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}. In \cite{HeloiseBeautification} used Zernike moment (FS3) and mentioned that order higher than 8 yield to little or no improvement. In our system we used much higher order to reach this saturation level. This is due to the level of details in the EL-DB and DL-DB sets. % achieved  HS-DB a
 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
	\caption{Feature Comparison on Hs-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure}

 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{paperImages/featuresLD.jpg.eps}
	\caption{Feature Comparison on LD-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure}

 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{results/FeatEL.eps}
	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure}

 The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}  
 \begin{figure}
	\centering
		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:LDtestFeatonly}
\end{figure}  
 \begin{figure}
	\centering
		\includegraphics[scale=0.8]{results/F1EL.eps}
	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeatonly}
\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testFeat.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllHS}
%\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testF1.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllMD}
%\end{figure}

>>>>

%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 50 to 60 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset.% In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from two different domains Electrical Engineering symbols (EL-DB) and logic design symbols (LD-DB). Data are collected from 7 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into 75\% training set and 25\% test set. Five different splits for the training and test data are generated from the dataset. %In worth noting that \cite{Oltmans07} gathered similar dataset and achieved 89\% on it   %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}

@@

 \begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
	\scalebox{0.7}{
		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
 Categories 	&ellipse	&heart	&trapezoid	&pentagon	&arch	&hexagon	&square	&triangle	&cube	&cylinder	&parallelogram	&moon	&callout\\ \hline 
ellipse	&150	&20	&0	&0	&0	&5	&1	&0	&0	&0	&0	&1	&0\\  \hline 
heart	&2	&130	&0	&1	&10	&1	&1	&0	&0	&0	&1	&4	&0\\  \hline 
trapezoid	&0	&1	&147	&5	&1	&0	&1	&3	&2	&0	&37	&0	&0\\  \hline 
pentagon	&0	&0	&1	&139	&0	&7	&0	&0	&0	&0	&1	&0	&0\\   \hline 
arch	&0	&0	&1	&1	&125	&0	&4	&0	&3	&0	&0	&15	&1\\   \hline 
hexagon	&0	&0	&0	&2	&0	&135	&0	&0	&0	&0	&0	&0	&0\\  \hline 
square	&0	&0	&0	&0	&0	&0	&138	&0	&0	&0	&0	&1	&0\\   \hline 
triangle	&0	&0	&1	&2	&5	&0	&1	&148	&0	&0	&4	&7	&0\\  \hline 
cube	&0	&0	&0	&1	&0	&0	&0	&0	&146	&5	&0	&0	&0\\   \hline 
cylinder	&0	&0	&0	&1	&0	&1	&0	&0	&0	&147	&0	&0	&0\\  \hline 
parallelogram	&0	&0	&0	&0	&0	&0	&2	&0	&2	&0	&107	&0	&0\\  \hline 
moon	&0	&1	&0	&0	&10	&2	&3	&0	&0	&0	&0	&125	&5\\   \hline 
callout	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&0	&145\\  \hline 
 		\end{tabular}
			}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix of HS-DB}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}

@@

\subsubsection{Features Comparisons}
\label{sec:featuresComparisions}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.  Figures \ref{fig:testFeaturesLDDB} and \ref{fig:ELtestFeaturesAll} show similar results on \textsl{EL-DB} and \textsl{LD-DB} dataset. The figures show that on both datasets \textbf{FS1} perform better than on \textsl{HS-DB}, this is a result that shapes in the \textsl{EL-DB} and \textsl{LD-DB} has more structural and conceptual differences than shapes in \textsl{HS-DB}. In \cite{HeloiseBeautification} used Zernike moment (FS3) and mentioned that order higher than 8 yield to little or no improvement. In our system we used much higher order to reach this saturation level. This is due to the level of details in the EL-DB and DL-DB sets.  Result shows that combining feature set FS4 and FS1 gives results compared to moments in both datasets \textsl{EL-DB} and \textsl{LD-DB}.    % achieved  HS-DB a
 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{paperImages/featuresdifference.jpg.eps}
	\caption{Feature Comparison on Hs-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure}

 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{paperImages/featuresLD.jpg.eps}
	\caption{Feature Comparison on LD-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesLDDB}
\end{figure}

 \begin{figure}
	\centering
		\includegraphics[scale=0.5]{results/FeatEL.eps}
	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeaturesAll}
\end{figure}

 The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}  
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
%	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:LDtestFeatonly}
%\end{figure}  
 \begin{figure}
	\centering
		\includegraphics[scale=0.8]{results/F1EL.eps}
	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeatonly}
\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testFeat.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllHS}
%\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testF1.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllMD}
%\end{figure}
